An Interpretable Machine Learning Framework for Modeling High-Resolution Spectroscopic Data\footnote{Open source code at \url{https://github.com/gully/blase}}

### 2022-09-23 paper/paper1/ms.tex
What would a Machine Learning practitioner call this?

\begin{document}
\shorttitle{Transfer learning with blas\'e}
\shortauthors{Gully \& Morley}
\title{An Interpretable Machine Learning Framework for Modeling High-Resolution Spectroscopic Data\footnote{Open source code at \url{https://github.com/gully/blase}}}

@@

@@

\noindent where the division indicates element-wise division of these arrays or ``vectors'' of flux values.  \edit1{For elementwise-multiplication, we will adopt the $\odot$ operator symbol to distinguish that the multiplication occurs elementwise between two ``vectors'' of spectral fluxes, as opposed to a dot product (scalar ouput) or matrix multiplication (matrix output).  This style of vector multiplication and division is standard in numerical numerical array manipulations such as those in \texttt{numpy}, \texttt{IDL}, \texttt{matlab}, \emph{etc.}}

@@

\emph{``A surrogate model, emulator, or discrepancy modeling''}-- The machine learning community is increasingly incorporating physics-based emulation into machine learning models.   \emph{Blas\'e} can be viewed as such a technique, often labeled as ``surrogate models''.

>>>>

\begin{document}
\shorttitle{\emph{blas\'e}: Interpretable Machine Learning for Spectroscopy}
\shortauthors{Gully-Santiago \& Morley}
\title{An Interpretable Machine Learning Framework for Modeling High-Resolution Spectroscopic Data\footnote{Open source code at \url{https://github.com/gully/blase}}}

@@

\edit1{Our first stage is to \emph{clone} a precomputed synthetic spectrum.  Cloning means to mimic, emulate, or otherwise approximate the appearance of a discretely sampled spectrum with some function or combination of functions.  The subject of the cloning shall be a noise-free spectrum generated from physics-based atmosphere models.  The purpose of cloning may appear superficial at first glance, but it serves as a necessary gateway for our ultimate data-model comparison goals.}

@@

\noindent where the division indicates element-wise division of these arrays or ``vectors'' of flux values.  \edit1{For elementwise-multiplication, we will adopt the $\odot$ operator symbol to distinguish that the multiplication occurs elementwise between two ``vectors'' of spectral fluxes, as opposed to a dot product (scalar ouput) or matrix multiplication (matrix output).  This style of vector multiplication and division is standard in numerical array manipulations such as those in \texttt{numpy}, \texttt{IDL}, \texttt{matlab}, \emph{etc.}}

@@

\emph{``Surrogate modeling / Simulation Intelligence''}-- \edit1{Our slightly off-label use of machine learning straddles a boundary of machine learning and conventional physics-based models.  The rapidly evolving field of machine learning may therefore classify the approach here using a few related terms.  The phrases surrogate modeling, emulation, discrepancy modeling, transfer learning, simulation intelligence, and physics-informed machine learning could describe some aspects of \emph{blas\'e}, depending on the ultimate application by the end-user.  For example, machine learning has been applied to physical chemistry models using a neural network as a cheaper-but-still-accurate ``surrogate'' for the real theory-based simulations. The realized speedups in quantum chemistry can approach 300,000$\times$ \citep{pmlr-v70-gilmer17a}. The key idea is that information from expensive physics-based simulations has been distilled by machine learning and can be applied to new examples.}

\emph{``Narrowing the Synthetic Gap'}-- \edit1{Callendar et al. describe the notion of a Synthetic Gap: the fact that real astronomical spectra to not resemble their machine learning clones...}
